---
title: "SDS315 HW3"
author: "Preston Bui"
date: "2024-02-19"
output: html_document
---

EID: ppb434

Github Repo: [https://github.com/PrestonBui/SDS315-HW3](https://github.com/PrestonBui/SDS315-HW3)

```{r message = FALSE, echo = FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(broom)
library(kableExtra)
library(stats)
```

# **Problem 1**
## Theory A: Competition Affects Gas Prices

**Claim:** Gas stations charge more if they lack direct competition in sight.

**Evidence:**

```{r echo = FALSE}
gas_prices <- read.csv("~/Downloads/gasprices.csv")

with_competitors_mean <- gas_prices %>% 
  filter(Competitors == 'Y') %>% 
  summarise(MeanPrice = mean(Price)) %>% 
  pull(MeanPrice)

without_competitors_mean <- gas_prices %>% 
  filter(Competitors == 'N') %>% 
  summarise(MeanPrice = mean(Price)) %>% 
  pull(MeanPrice)

mean_diff <- without_competitors_mean - with_competitors_mean

ci <- t.test(gas_prices$Price ~ gas_prices$Competitors)$conf.int
```

- Mean price with competitors: `r round(with_competitors_mean, 3)` USD/gallon

- Mean price without competitors: `r round(without_competitors_mean, 3)` USD/gallon

- Mean difference: `r round(mean_diff, 3)` USD/gallon

- 95% confidence interval for the mean difference: `r round(ci[1], 3)` to `r round(ci[2], 3)`

**Conclusion: ** The data supports the claim. Gas stations without direct competition in sight tend to charge more, with a statistically significant mean difference in price.

## Theory B: Wealthier Areas Pay More

**Claim:** The richer the area, the higher the gas prices.

**Evidence:**

```{r echo = FALSE}
linear_model <- lm(Price ~ Income, data = gas_prices)
summary(linear_model)$coefficients
```

**Conclusion:** The regression analysis reveals a positive and statistically significant relationship between area income and gas prices, with a coefficient of `r format(1.248e-06, scientific = TRUE)` for Income. This implies that as the median income in an area increases, gas prices also increase. The p-value of `r format(4.116e-05, scientific = TRUE)` confirms the statistical significance of this finding. Therefore, we conclude that the data supports the claim that wealthier areas have higher gas prices.

## Theory C: Stoplights Increase Prices

**Claim:** Gas stations at stoplights charge more.

**Evidence:**
```{r echo = FALSE}
t_test_c <- t.test(Price ~ Stoplight, data = gas_prices)
t_test_c
```

**Conclusion:** The Welch Two Sample t-test indicates that the difference in mean gas prices between stations with and without stoplights is not statistically significant (p-value = 0.8521). The 95% confidence interval for the mean price difference ranges from -0.0318 to 0.0385 USD per gallon, which includes zero. Therefore, we conclude that the data does not support the claim that gas stations at stoplights charge more; the evidence suggests that stoplights do not have a noticeable effect on gas pricing.


## Theory D: Highway Access Increases Prices

**Claim:** Highway Access Increases Prices

**Evidence:**
```{r echo = FALSE}
t_test_d <- t.test(Price ~ Highway, data = gas_prices)
t_test_d

```

**Conclusion:** The Welch Two Sample t-test provides evidence that there is a statistically significant difference in gas prices between stations with and without highway access (p-value = 0.0188). The negative t-value suggests that gas stations with highway access actually have lower prices on average, contrary to the claim. The 95% confidence interval for the mean difference in price ranges from approximately -0.0834 to -0.0080 USD per gallon, further supporting that highway access does not lead to increased prices. Thus, we conclude that the data contradicts the claim; stations with highway access do not charge more, but rather less.


## Theory E: Shell Charges More

**Claim:** Shell charges more than all other non-Shell brands.

**Evidence:**
```{r echo = FALSE}
gas_prices$IsShell <- as.factor(gas_prices$Brand == "Shell")
t_test_e <- t.test(Price ~ IsShell, data = gas_prices)
t_test_e
```

**Conclusion:** The Welch Two Sample t-test results indicate that the mean price for Shell branded gas stations is not significantly higher than that of non-Shell brands (p-value = 0.1603). The confidence interval for the mean difference in price ranges from approximately -0.0661 to 0.0113 USD per gallon, which includes zero. This suggests that there is no statistically significant evidence to support the claim that Shell charges more than other brands. The means of the two groups are statistically indistinguishable, leading to the conclusion that Shell does not consistently charge more for gas compared to other brands.


# **Problem 2**
## Part A: Average Mileage of 2011 S-Class 63 AMGs
```{r echo = FALSE}
sclass <- read.csv("~/Downloads/sclass.csv")
filtered_a <- sclass %>%
  filter(year == 2011, trim == "63 AMG")


bootstrap_a <- replicate(1000, sample(filtered_a$mileage, size = nrow(filtered_a), replace = TRUE)) %>%
  apply(2, mean)

mileage_ci <- quantile(bootstrap_a, probs = c(0.025, 0.975))
```

The 95% confidence interval for the average mileage of 2011 S-Class 63 AMGs is approximately `r format(mileage_ci[1], big.mark = ",")` to `r format(mileage_ci[2], big.mark = ",")` miles.

## Part B: Proportion of Black 2014 S-Class 550s
```{r echo = FALSE}
filtered_b <- sclass %>%
  filter(year == 2014, trim == "550") %>%
  mutate(isBlack = as.integer(color == "Black"))

bootstrap_b <- replicate(1000, sample(filtered_b$isBlack, size = nrow(filtered_b), replace = TRUE)) %>%
  apply(2, mean)

black_ci <- quantile(bootstrap_b, probs = c(0.025, 0.975))

```

The 95% confidence interval for the proportion of black 2014 S-Class 550s is approximately `r round(black_ci[1] * 100, 2)`% to `r round(black_ci[2] * 100, 2)`%.

# **Problem 3**
## Part A
**Question: ** Which show makes people happier: "Living with Ed" or "My Name is Earl"?

**Approach:** We filtered the dataset for these two shows and used t.test() to compute the 95% confidence interval for the difference in mean "Q1_Happy" scores.

```{r echo = FALSE}
survey_data <- read.csv("~/Downloads/nbc_pilotsurvey.csv")

filtered_a <- survey_data %>% filter(Show %in% c("Living with Ed", "My Name is Earl"))

ci_a <- t.test(Q1_Happy ~ Show, data = filtered_a, conf.level = 0.95)$conf.int
```

**Results: ** The 95% confidence interval for the difference in mean happiness scores between "Living with Ed" and "My Name is Earl" is `r round(ci_a[1], 2)` to `r round(ci_a[2], 2)`.

**Conclusion ** The 95% confidence interval for the difference in mean happiness scores between "Living with Ed" and "My Name is Earl" ranges from -0.1 to 0.4. Since this interval includes zero, it suggests that there is not a statistically significant difference in happiness scores between the two shows. Thus, we do not have enough evidence to conclude that either "Living with Ed" or "My Name is Earl" consistently makes viewers happier.

## Part B
**Question: ** Which reality/contest show made people feel more annoyed?

**Approach: ** We computed the 95% confidence interval for the difference in mean "Q1_Annoyed" scores.

```{r echo = FALSE}
filtered_b <- survey_data %>% filter(Show %in% c("The Biggest Loser", "The Apprentice: Los Angeles"))
ci_b <- t.test(Q1_Annoyed ~ Show, data = filtered_b, conf.level = 0.95)$conf.int
```

**Results: ** The 95% confidence interval for the difference in mean annoyance scores is `r round(ci_b[1], 2)` to `r round(ci_b[2], 2)`.

**Conclusion ** The 95% confidence interval for the difference in mean annoyance scores is 0.02 to 0.52. Because this interval does not include zero, it suggests that there is a statistically significant difference in annoyance scores between "The Biggest Loser" and "The Apprentice: Los Angeles." However, without the direction of the difference (which show has the higher mean), we cannot conclude which show made people feel more annoyed. You will need to check which show has the higher mean annoyance score to make a definitive conclusion.

## Part C
**Question: ** What proportion of viewers found "Dancing with the Stars" confusing?

**Approach: ** We computed a large-sample 95% confidence interval for the proportion of "Q2_Confusing" scores of 4 or greater.

```{r echo = FALSE}
filtered_c <- survey_data %>% filter(Show == "Dancing with the Stars", Q2_Confusing >= 4)
total_respondents <- nrow(survey_data %>% filter(Show == "Dancing with the Stars"))
ci_c <- prop.test(x = nrow(filtered_c), n = total_respondents, conf.level = 0.95)$conf.int
```

**Results: ** The 95% confidence interval for the proportion of viewers who found "Dancing with the Stars" confusing is `r round(ci_c[1], 4)` to `r round(ci_c[2], 4)`.

**Conclusion ** The 95% confidence interval for the proportion of viewers who found "Dancing with the Stars" confusing is 0.0445 to 0.1289. This indicates that between 4.45% and 12.89% of viewers found the show confusing. While this suggests that the majority of viewers were not confused by the show, there is a notable minority that was. This could be an area for the show's producers to investigate further to see if the show's format can be made clearer.

# **Problem 4**
**Question: ** Is there a difference in the revenue ratio between DMAs that had paid search advertising paused and those that did not?

**Approach: ** We will calculate the revenue ratio for each DMA, perform a bootstrap analysis with 10,000 simulations to estimate the difference in revenue ratios between treatment and control groups, and then compute a 95% confidence interval for this difference.

```{r calculate-ratio, echo=FALSE}
ebay_data <- read.csv("~/Downloads/ebay.csv")

ebay_data <- ebay_data %>%
  mutate(revenue_ratio = rev_after / rev_before)

bootstrap_diff <- function(data){
  sample_data <- data %>% 
    sample_frac(size = 1, replace = TRUE)
  
  treatment_mean <- mean(sample_data$revenue_ratio[sample_data$adwords_pause == 1])
  control_mean <- mean(sample_data$revenue_ratio[sample_data$adwords_pause == 0])
  
  treatment_mean - control_mean
}

bootstrap_results <- replicate(10000, bootstrap_diff(ebay_data))

ci <- quantile(bootstrap_results, probs = c(0.025, 0.975))
ci
```

**Results: ** The bootstrap analysis yields a 95% confidence interval for the difference in revenue ratios between the treatment and control groups, ranging from approximately -0.0905 to -0.0139.

**Conclusion: ** The 95% confidence interval suggests that the revenue ratio is consistently lower in the treatment group (where paid search advertising was paused) compared to the control group. Since the confidence interval does not include zero and indicates a decrease in revenue ratio, we conclude that paid search advertising on Google likely contributes positively to eBay's revenue. The absence of these ads appears to have a statistically significant negative impact on the revenue ratios in the treatment DMAs.

